### 2024-11-04 paper/paper.qmd
fixed error in string

---
title: "Prediction of 2024 US election ..."
author: 
  - Colin Sihan Yang
  - Lexun Yu
  - Siddharth Gowda
thanks: "Code and data are available at: [https://github.com/yulexun/uselection](https://github.com/yulexun/uselection)."
date: today
date-format: long
abstract: "We forecast the winner of the 2024 US presidential election using “poll-of-polls” by building a linear model."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(here)
library(tibble)
library(knitr)
library(brms)
library(rstanarm)
library(gridExtra)
library(Matrix)
library(rstan)
library(brms)

```

```{r}
#| echo: false
#| warning: false
cleaned_data = read_csv(here("data/02-analysis_data/cleaned_data.csv"))
```

# Introduction {#sec-intro}
Election result forecasting has become an essential tool for analysts in political science and the public to predict the outcome of democratic process, such as the presidential election in the United States. Traditionally, individual polls have been used as a snapshot of voter sentiment, but they only reflect temporary changes in the performance of contestants, instead of a precise estimation of the election result. As discussed by @pasek and @blumenthal, the aggregation of multiple polls, or "poll-of-polls," has become a popular technique to reduce individual survey errors and provide more accurate election forecasts. However, the traditional poll aggregation does not reflect dynamics of an election, especially with real-time changes and the introduction of new data. This creates a gap for a more adaptable model to predict the election result based on both polling data and additional variables, such as historical data and economic indicators.

This paper fills the gap by building a hybrid election forecasting model following the strategies mentioned by @pasek. As @pasek described in their article, aggregation involves determining which surveys are worth including, as well as selecting, combining and averaging results from multiple polls to reduce individual biases and errors. Prediction modeling adds other data to the model that predicts election outcomes based on current dynamics. Hybrid models like the Bayesian approach incorporates prior beliefs based on historical data or expert knowledge and new evidence like economic updates to dynamically adjust the forecast as the campaign progresses.

In this paper, we aim to predict the 2024 us election result with the hybrid election forcasting model. We incorporate aggregation by filtering the polls on @fivethirtyeight by numeric grade that indicates pollster’s reliability, prediction that incorporates social and economic indicators including unemployment rates and abortion rates, and hybrid approaches that leverages Bayesian techniques which combines historical data such as the 2016 election data, allowing for a dynamic prediction of the U.S. presidential election.

The estimand for this research paper is the predicted support percentages for Kamala Harris and Donald Trump. The prediction is based on quantifying various polling factors, including sample size, poll scores, and transparency scores, which are used as predictors.

The results of this model indicate a more stable and accurate forecast compared to traditional aggregation methods alone, [update this …]

The remainder of this paper is structured as follows: [update this …]

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

# FiveThirtyEight Licenses
[FiveThirtyEight's data sets](https://github.com/fivethirtyeight/data/tree/master/polls) are used and modified by us under the [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/).

# Our Prediction

```{r}
#| echo: false
#| warning: false
# Split data into training and test sets

set.seed(123)

train_data_harris <- read_csv(here("data/02-analysis_data/train_data_harris.csv"))
test_data_harris <- read_csv(here("data/02-analysis_data/test_data_harris.csv"))
bayesian_model_train_harris <- readRDS(here("models/bayesian_model_train_harris.rds"))

train_data_trump <- read_csv(here("data/02-analysis_data/train_data_trump.csv"))
test_data_trump <- read_csv(here("data/02-analysis_data/test_data_trump.csv"))
bayesian_model_train_trump <- readRDS(here("models/bayesian_model_train_trump.rds"))

relevant_states <- c("Arizona", "Nevada", "Georgia", "Pennsylvania", "Michigan", "Minnesota",
                     "Wisconsin", "Florida", "Texas", "Maine CD-2", "Nebraska CD-2", 
                     "New Hampshire", "Ohio", "Virginia", "North Carolina", "Iowa")

# Initialize an empty tibble to store results
swing_state_predictions <- tibble(
  state = character(),
  harris_predicted_pct = numeric(),
  trump_predicted_pct = numeric(),
  winner = character()
)

for(state in relevant_states) {
  #print(state, state %in% unique(test_data_trump), state %in% unique(test_data_harris))
  test_data_state_harris <- test_data_harris[test_data_harris$state == state, ]
  test_data_state_trump <- test_data_trump[test_data_trump$state == state, ]

  predictions_state_harris <- posterior_predict(bayesian_model_train_harris, newdata = test_data_state_harris)
  predictions_state_trump <- posterior_predict(bayesian_model_train_trump, newdata = test_data_state_trump)

  avg_predicted_pct_state_harris <- mean(predictions_state_harris)
  avg_predicted_pct_state_trump <- mean(predictions_state_trump)

  # Add a new row to the results tibble
  new_row <- tibble(
    state = state,
    harris_predicted_pct = avg_predicted_pct_state_harris,
    trump_predicted_pct = avg_predicted_pct_state_trump,
    winner = ifelse(trump_predicted_pct > harris_predicted_pct, "Trump", "Harris")
  )
  # print(state)
  # print(avg_predicted_pct_state_harris)
  # print(avg_predicted_pct_state_trump)
  # 
  swing_state_predictions <- bind_rows(swing_state_predictions, new_row)
}

swing_state_predictions <- swing_state_predictions %>% arrange(state) %>% unique()

# View the results
swing_state_predictions

```

```{r}
#| label: fig-currentstateofelection
#| fig-cap: 2024 U.S. Presidential Election State Forecast Map
#| echo: false
#| warning: false

# Define your colors for each state (replace these colors with your desired ones)
state_colors_legend <- c(
  "Republican" = "red",      # Dark Red
  "Democrat" = "blue",        # Dark Blue
  "Split By District" = "gray"
)

purple = "purple"
light_blue = "green"
light_red = "orange"

state_colors <- c(
  alabama = "red",
  arizona = ifelse((swing_state_predictions %>% 
                        filter(state == "Arizona") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  arkansas = "red",
  california = "blue",
  colorado = "blue",
  connecticut = "blue",
  delaware = "blue",
  `district of columbia` = "blue",
  florida = ifelse((swing_state_predictions %>% 
                        filter(state == "Florida") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  georgia = ifelse((swing_state_predictions %>% 
                        filter(state == "Georgia") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  idaho = "red",
  illinois = "blue",
  indiana = "red",
  iowa = ifelse((swing_state_predictions %>% 
                        filter(state == "Iowa") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  kansas = "red",
  kentucky = "red",
  louisiana = "red",
  maine = "gray",
  maryland = "blue",
  massachusetts = "blue",
  michigan = ifelse((swing_state_predictions %>% 
                        filter(state == "Michigan") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  minnesota = ifelse((swing_state_predictions %>% 
                        filter(state == "Minnesota") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  mississippi = "red",
  missouri = "red",
  montana = "red",
  nebraska = "gray",
  nevada = ifelse((swing_state_predictions %>% 
                        filter(state == "Nevada") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  `new hampshire` = ifelse((swing_state_predictions %>% 
                        filter(state == "New Hampshire") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  `new jersey` = "blue",
  `new mexico` = "blue",
  `new york` = "blue",
  `north carolina` = ifelse((swing_state_predictions %>% 
                        filter(state == "North Carolina") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  `north dakota` = "red",
  ohio = ifelse((swing_state_predictions %>% 
                        filter(state == "Ohio") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  oklahoma = "red",
  oregon = "blue",
  pennsylvania = ifelse((swing_state_predictions %>% 
                        filter(state == "Pennsylvania") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  `rhode island` = "blue",
  `south carolina` = "red",
  `south dakota` = "red",
  tennessee = "red",
  texas = ifelse((swing_state_predictions %>% 
                        filter(state == "Texas") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  utah = "red",
  vermont = "blue",
  virginia = ifelse((swing_state_predictions %>% 
                        filter(state == "Virginia") %>% 
                        select(winner)) == "Trump", "red", "blue"),
  washington = "blue",
  `west virginia` = "red",
  wisconsin = ifelse((swing_state_predictions %>% 
                        filter(state == "Wisconsin") %>% 
                        select(winner)) == "Trump", "red", "blue"), 
  wyoming = "red")

# Load the states map data
states_map <- map_data(map = "state")

state_colors <- setNames(state_colors, tolower(names(state_colors)))

color_to_status <- setNames(names(state_colors_legend), state_colors_legend)

states_map <- states_map %>%
  mutate(status = color_to_status[state_colors[region]])

plotclr <- c("red", gray, "blue")
legend_description <- c(
  "Republican",
  "Democrat",
  "Split by District"
)

ggplot(states_map, aes(x = long, y = lat, group = group, fill = status)) +
  geom_polygon(color = "white", size = 0.2) +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  scale_fill_manual(
    values = state_colors_legend,
    name = "Election Status"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10),
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank()
  )

```
Maine CD-2	47.40416	49.31394	Trump
Nebraska CD-2	49.90928	42.17853	Harris

\newpage

# References